---
title: "Finding Density Peaks"
author: "Anton Holm"
date: '2022-03-19'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Rfast)
library(tidyverse)

```

Small dataset
```{r}
#### START ####
source("..//Full_Method_Updated.R")
source("..//Generate_GMM_Data.R")
#Load Data
daata = readRDS("..//DE (250K genes, varying cellsize + cellshape)//(DATA) n = 57.000, c = 250, tissue = 400, VS.rdata")
#daata = Generate_GMM_Data(clusters = 6000, tissue_x = 500, tissue_y = 500, Cellwidth = 11.6, Mode = 3)
mode_distances <- dista(daata$Mu, daata$Mu)
diag(mode_distances) <- NA
avg_mode_dist <- mean(apply(mode_distances, 1, min, na.rm = TRUE))
avgPeakDist <- avg_mode_dist
cluster_peaks <- findClusterPeaks(daata$Coordinates, avgPeakDist = avgPeakDist, k = 7, estimator = 2)

ggplot(data = NULL) +
  geom_point(aes(x = daata$Coordinates[,1], y = daata$Coordinates[,2], col = daata$Samples)) +
  geom_point(aes(x = daata$Mu[,1], daata$Mu[,2]), color = "red", size = 3) +
  geom_point(aes(x = daata$Coordinates[cluster_peaks$ClusterPeaks,1], y = daata$Coordinates[cluster_peaks$ClusterPeaks,2])) +
  guides(col = "none") +
  geom_hline(yintercept = ymin + 0:ygrids*grid_size) +
  geom_vline(xintercept = xmin + 0:xgrids*grid_size)
```


















Simulated GMM
-----------------------------------------------------------------
osmFish data


```{r}
library(loomR)
osmData <- readRDS("..//SSAM Reconstructed//gene coordinates osmFISH DF.rdata")

osmPeaks <- findClusterPeaks(cbind(osmData$V1, osmData$V2), avgPeakDist = 8, k = 7, estimator = 2)


```


















osmFish data
------------------------------------------------------------------
Notes

Last Last time:
 - Fixed automatic way to find threshold
  - used kernel regression here
 - Fixed loop to loop over grids
 
Last Time:
 - Made the threshold finding for each grid instead of one at the end
 - The loop now calculates correct deltas, exactly the same as for the method of doing entire images at once.
 - I get all the same cluster centers if I dont do automatic and do all at the end. Only miss those at the boundries as we should.
 - Distance between consecutive threshold lines increase as we move up the graph.
 - I only look at small grids with at least 6 points. Is this ok? Should we increase this threshold?
 - I let the right most part of the monotonic regression line go up to L/4 and it works
Last time 3: 
- Calculating Deltas by grids works perfect
- Finding number of clusters manually works perfect
- All calculations works great.

Next time:
- Final problem: I get R sessions aborted when running with large data
- Clustering

Look at clustering from fast search paper (look at closest neighbor, then second and so on.)
