---
title: "Density Estimation"
author: "Anton Holm"
date: '2022-02-14'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(MASS)
library(tidyverse)
library(spatstat)
library(dbscan)
library(png)
library(mvtnorm)
```

Sample from GMM similar to OSMfish data.
```{r}
set.seed(931031)

clusters = 6000
avg_genes = 200
cellwidth = 11.6
  ## A function to create a dataframe consisting of a GMM sample
  ## comparable to some in-situ sample
  
  # :Param clusters: Decides the number of cells
  # :Param avg_genes: Average number of genes per cell
  # :Param cellwidth: Used as component variance in covariance matrix (cell width in um)

  #600 different mean vectors (600 different gaussians in GMM)
  mu <-cbind(runif(clusters, 0, 2000), runif(clusters, 0, 3000))
  #Constant covariance matrix. 11.6um per cell
  Cov <- matrix(c(cellwidth, 0, 0, cellwidth), nrow = 2, ncol = 2)
  #Decide which gaussian to sample from (same probabilities)
  GMM_sample <- sample.int(clusters, clusters*avg_genes, replace = TRUE) %>% 
    table()
  # Create Sample from GMM
  gene_coord <- matrix(NA, ncol = 2, nrow = 0)
  true_dens <- matrix(NA, ncol = 1, nrow = 0)
  for (i in 1:length(GMM_sample)){
    k <- as.numeric(dimnames(GMM_sample)[[1]][i])
    mu1 <- mu[k, ]
    point <- mvrnorm(n = GMM_sample[i], mu = mu1, Sigma = Cov)
    gene_coord <- rbind(gene_coord, point)
    true_dens <- c(true_dens, dmvnorm(point, mean = mu1, sigma = Cov))
  }
```


Two density estimations (Kraskov and KDE). Quite similar results and ranks are more similar at sparse areas. why?

Density estimation and comparison (Kraskov)
```{r}
###################### Kraskov Density Estimation and comparison ###################################

# Estimation of density based on Kraskov Paper 
kraskov_estim <- function(x, N, k, d){
  exp(digamma(k) - digamma(N) - d*log(2*nndist(x, k = k)) - log(pi^(d/2)/gamma(1 + d/2)))
}

kraskov_dens <- kraskov_estim(gene_coord, N = nrow(gene_coord), k = 7, d = 2)

cor.test(rank(kraskov_dens), rank(true_dens), method = 'spearman')

rank_comparison <- ggplot(as_tibble(cbind(rank(-kraskov_dens), rank(-true_dens))), aes(x = V1, y = V2)) +
    geom_density_2d_filled(contour_var = 'ndensity') +
  labs(title = "Kraskov vs True Rank: Clusters = 6000, N = 1.2M, Var = 11.6",
       x = "Kraskov Rank",
       y = "True Rank")
ggsave(filename = "rank_comparison.png")
```


Density estimation and comparison (KDE: Gaussian Kernel with bandwidth 2.5)

```{r}
kernel <- function(x, h){
  1/(2*pi*h^2) * exp(-1/2 * (x/h)^2)
}

#Calculate euclidean distance to all k-nearest neighbors (list with each gene-type in the list)
knn_distance <- kNNdist(gene_coord, k = 7, all = TRUE)

kde_dens <- kernel(knn_distance, h = 2.5) %>% 
  rowSums()

cor.test(rank(kde_dens),rank(true_dens), method = "spearman")

kde_vs_true <- ggplot(as_tibble(cbind(rank(-kde_dens), rank(-true_dens))), aes(x = V1, y = V2)) +
    geom_density_2d_filled(contour_var = 'ndensity') +
  labs(title = "KDE vs True Rank: Clusters = 6000, N = 1.2M, Var = 11.6",
       x = "KDE Rank",
       y = "True Rank")
ggsave(filename = "rank_comparison_kde.png")
```



Spearman vs N

Starts to coverge at around N = 500000 (this is for 6000 clusters) i.e. around 100 samples per cell on avg

```{r}
n <- c(10000, 50000, 100000, 500000, 1000000, 3000000)
spearman_ranks <- c()
clusters = 6000

for (j in 1:length(n)){
  #Decide which gaussian to sample from (same probabilities)
  GMM_sample2 <- sample.int(clusters, n[j], replace = TRUE) %>% 
    table()
  # Create Sample from GMM
  gene_coord2 <- matrix(NA, ncol = 2, nrow = 0)
  true_dens2 <- matrix(NA, ncol = 1, nrow = 0)
  kraskov_dens2 <- matrix(NA, ncol = 1, nrow = 0)
  for (i in 1:length(GMM_sample2)){
    k2 <- as.numeric(dimnames(GMM_sample2)[[1]][i])
    mu2 <- mu[k2, ]
    point2 <- rmvnorm(n = GMM_sample2[i], mean = mu2, sigma = Cov)
    gene_coord2 <- rbind(gene_coord2, point2)
    true_dens2 <- c(true_dens2, dmvnorm(point2, mean = mu2, sigma = Cov))
  }

  kraskov_dens2 <- kraskov_estim(gene_coord2, N = n[j], k = 7, d = 2)
  spearman_ranks[j] <- cor.test(rank(true_dens2), rank(kraskov_dens2), method = "spearman", ties.method = "average")$estimate
}

ggplot(as_tibble(cbind(n, spearman_ranks)), aes(x = n, y = spearman_ranks)) +
  geom_point() +
  labs(title = "Spearman rank vs N")
```











Some Testing



Alternating size and shape of cells.
Similar results between KDE and Kraskov still
```{r}
clusters = 6000
avg_genes = 200
cellwidth = 11.6
  ## A function to create a dataframe consisting of a GMM sample
  ## comparable to some in-situ sample
  
  # :Param clusters: Decides the number of cells
  # :Param avg_genes: Average number of genes per cell
  # :Param cellwidth: Used as component variance in covariance matrix (cell width in um)

  gene_coord_diff <- matrix(NA, ncol = 2, nrow = 0)
  true_dens_diff <- matrix(NA, ncol = 1, nrow = 0)
  for (i in 1:length(GMM_sample)){
    k <- as.numeric(dimnames(GMM_sample)[[1]][i])
    mu1 <- mu[k, ]
    eps1 <- rnorm(1, 0, 1)
    eps2 <- rnorm(1, 0, 1)
    Cov_diff <- matrix(c(sqrt(cellwidth) + eps1, eps2, eps2, sqrt(cellwidth) + eps1), nrow = 2, ncol = 2)
    point_diff <- mvrnorm(n = GMM_sample[i], mu = mu1, Sigma = t(Cov_diff) %*% Cov_diff)
    gene_coord_diff <- rbind(gene_coord_diff, point_diff)
    true_dens_diff <- c(true_dens_diff, dmvnorm(point_diff, mean = mu1, sigma = Cov_diff))
  }
  
  
# Kraskov
kraskov_dens_diff <- kraskov_estim(gene_coord_diff, N = nrow(gene_coord_diff), k = 7, d = 2)

kras_spearman_diff <- cor.test(rank(kraskov_dens_diff), rank(true_dens_diff), method = 'spearman')$estimate

rank_comparison_diff_kras <- ggplot(as_tibble(cbind(rank(-kraskov_dens_diff), rank(-true_dens_diff))), aes(x = V1, y = V2)) +
    geom_density_2d_filled(contour_var = 'ndensity') +
  labs(title = "Kraskov vs True Rank: Clusters = 6000, N = 1.2M, Var = 11.6 (varying cellsize and shape)",
       x = "Kraskov Rank",
       y = "True Rank")
ggsave(filename = "rank_comparison_diff_kras.png")

# KDE
knn_distance_diff <- kNNdist(gene_coord_diff, k = 7, all = TRUE)

kde_dens_diff <- kernel(knn_distance_diff, h = 2.5) %>% 
  rowSums()

kde_spearman_diff <- cor.test(rank(-kde_dens_diff),rank(-true_dens_diff), method = "spearman")$estimate

rank_comparison_diff_kde <- ggplot(as_tibble(cbind(rank(-kde_dens_diff), rank(-true_dens_diff))), aes(x = V1, y = V2)) +
    geom_density_2d_filled(contour_var = 'ndensity') +
  labs(title = "KDE vs True Rank: Clusters = 6000, N = 1.2M, Var = 11.6 (varying cellsize and shape)",
       x = "KDE Rank",
       y = "True Rank")
ggsave(filename = "rank_comparison_diff_kde.png")

cbind(kras = kras_spearman_diff, kde = kde_spearman_diff)
```





Make some sparse areas

Similar results between kde and kras. some of the lowest true densities are given highest densities using both KDE and kraskov.

```{r}
df <- gene_coord_diff %>% 
  as_tibble() %>% 
  rename("x" = V1, "y" = V2) %>% 
  mutate(true_dens = true_dens_diff)

df1 <- df %>% 
  filter(x < 700 & y < 1000) %>% 
  sample_frac(0.8)

df2 <- df %>% 
  filter(x > 1400 & y < 2000 & y > 1000) %>% 
  sample_frac(0.9)

df3 <- df %>% 
  filter(x < 700 & y > 2000) %>% 
  sample_frac(0.8)

df4 <- df %>% 
  filter(x > 1400 & y > 2000)


sparse_data <- anti_join(df, df1, by = c("x", "y")) %>% 
  anti_join(df2, by = c("x", "y")) %>% 
  anti_join(df3, by = c("x", "y")) %>% 
  anti_join(df4, by = c("x", "y")) 

sparse_data_plot <- sparse_data %>% 
  ggplot(aes(x, y)) +
    geom_density_2d_filled(contour_var = 'ndensity')
ggsave(filename = "Sparse_data.png")


#Kraskov
kras_sparse <- kraskov_estim(as.matrix(sparse_data[,1:2]), N = nrow(sparse_data), k = 7, d = 2)
kras_spearman_sparse <- cor.test(rank(-kras_sparse), rank(-sparse_data$true_dens), method = 'spearman')$estimate
  
rank_comparison_sparse_kras <- ggplot(as_tibble(cbind(rank(-kras_sparse), rank(-sparse_data$true_dens))), aes(x = V1, y = V2)) +
    geom_density_2d_filled(contour_var = 'ndensity') +
  labs(title = "Kraskov vs True Rank: Clusters = 6000, N = 1.2M, Var = 11.6 (sparse)",
       x = "Kraskov Rank",
       y = "True Rank")
ggsave(filename = "rank_comparison_sparse_kraskov.png")


# KDE
knn_distance_sparse <- kNNdist(as.matrix(sparse_data[,1:2]), k = 7, all = TRUE)

kde_dens_sparse <- kernel(knn_distance_sparse, h = 2.5) %>% 
  rowSums()

kde_spearman_sparse <- cor.test(rank(kde_dens_sparse),rank(sparse_data$true_dens), method = "spearman")$estimate

rank_comparison_diff_kde <- ggplot(as_tibble(cbind(rank(-kde_dens_sparse), rank(-sparse_data$true_dens))), aes(x = V1, y = V2)) +
    geom_density_2d_filled(contour_var = 'ndensity') +
  labs(title = "KDE vs True Rank: Clusters = 6000, N = 1.2M, Var = 11.6 (sparse)",
       x = "KDE Rank",
       y = "True Rank")
ggsave(filename = "rank_comparison_sparse_KDE.png")

cbind(kras = kras_spearman_sparse, kde = kde_spearman_sparse)
```

Check abnomaly in highest and lowest ranks

Seems again like the estimations are fine with sparse areas but have problems with dense areas
```{r}
outliers <- as_tibble(cbind(kras = rank(-kras_sparse), true = rank(-sparse_data$true_dens))) %>% 
  rowid_to_column() %>% 
  filter(true > 600000 & kras < 200000)

sparse_data %>% 
  rowid_to_column() %>% 
  filter(rowid %in% outliers$rowid) %>% 
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  labs(title = "Points with high kraskov density and low true density")
```




Plots over kde vs true density and kraskov vs true density

```{r}

ggplot(as_tibble(cbind(kde_dens, true_dens)) %>% 
                   sample_frac(0.1), aes(x = kde_dens, y = true_dens)) + 
  geom_point() +
  xlim(0,0.25) +
  ylim(0,0.25)


ggplot(as_tibble(cbind(kraskov_dens, true_dens)) %>% 
                   sample_frac(0.1), aes(x = kraskov_dens, y = true_dens)) + 
  geom_point() +
  xlim(0, 0.014) +
  ylim(0, 0.014)


```





Check if kras can find number of modes


Check bias and variance of kraskov vs KDE,

Strange that we have good rank comparison on sparse areas and worse the denser it gets. Why is that?

How do we decide on k in density estimation?

How do we estimate number of modes/peaks?